---
title: "Data Acquisition"
date: 2021-05-09
author: Barkschat Jens, Beyerle Gregor
output:
  html_notebook:
    toc: TRUE
---

## Finding Possible Ticker Symbols

As we're looking at S&P 500 data and plan to cluster the different companies
with regard to their performance (and crisis performance) we'll try to get as many
companies, that have been listed in the S&P 500 and exist to this day.


```{r}
library(tidyverse)
library(tsibble)
library(feasts)
library(janitor)
library(qrmdata)
data("SP500_const")
```

We take all ticker symbols that have been in the S&P 500 since 1996. We use the
tracking data supplied in this [repository](https://github.com/fja05680/sp500).

```{r}
s_and_p_movements <- read_csv("data/S&P_500_Historical_Components_and_Changes_2021_25_04.csv")
```

For all these years we'll collect all symbols and keep the unique ones for further processing.

```{r}
unique_ticker_symbols <- unique(unlist(strsplit(s_and_p_movements$tickers[1:10], split = ",")))
length(unique_ticker_symbols)
```

## Check Ticker Symbol Availability

We use the `TTR` package to find out which of those symbols are still commonly
available. S&P 500 indizes are usually commonly traded but companies, that do not exist
anymore will surely not be listed anymore.

```{r}
current_common_exchange_symbols <- TTR::stockSymbols()
length(current_common_exchange_symbols$Symbol)
```

If our symbols aren't available anymore we will throw them out.

```{r}
is_currently_listed <- unique_ticker_symbols %in% current_common_exchange_symbols$Symbol
still_listed_symbols <- unique_ticker_symbols[is_currently_listed]
length(still_listed_symbols)
```

```{r}
still_listed_symbols
```

## Collect Ticker Data

We will use the remaining 276 symbols and lookup the daily stock data for them.
In the background this will call the Yahoo stock information service.
We include all datapoints up to 2021-05-09 and try to go back to 1991 so we'd have
still a year of data before the dotcom bubble burst.

```{r}
start_date <- as.Date("1999-01-01")
end_date <- as.Date("2021-05-09")
freqency <- "daily"

lookup_output <- BatchGetSymbols::BatchGetSymbols(tickers = still_listed_symbols,
                                                  first.date = start_date,
                                                  last.date = end_date,
                                                  freq.data = freqency,
                                                  cache.folder = file.path(tempdir(), "BGS_Cache"))
```

Because we can't sensibly impute data for these financial time series we'll only
include symbols for which there are all price data available.

```{r}
good_to_go_symbols <- lookup_output$df.control %>%
  filter(perc.benchmark.dates == 1.0) %>%
  (function (x) { x$ticker })

ticker_values <- lookup_output$df.tickers %>%
  filter(ticker %in% good_to_go_symbols) %>%
  as_tsibble(key = ticker, index = ref.date)

ticker_values
```

```{r}
good_to_go_meta <- current_common_exchange_symbols %>%
  filter(Symbol %in% good_to_go_symbols)

good_to_go_meta
```

```{r}
additional_meta <- SP500_const_info %>%
  filter(Ticker %in% good_to_go_symbols) %>%
  rename(Symbol = Ticker)

additional_meta
```

```{r}
complete_meta <- good_to_go_meta %>%
  remove_empty(which = c("cols")) %>%
  select(-c(ACT.Symbol, CQS.Symbol, NASDAQ.Symbol, Test.Issue, ETF)) %>%
  left_join(additional_meta, by = "Symbol")

complete_meta
```

## Quick Sanity Check

At this point we have complete daily time series for 216 ticker symbols.
We can plot the daily closing prices for one symbol (AAPL for Apple) to check if
the data we pulled down makes sense.

```{r}
ticker_values %>%
  filter(ticker == "AAPL") %>%
  autoplot(price.close)
```

## Persisting The Data

When we're happy with the data we can save it for later processing.

```{r}
ticker_values %>% write_csv("data/ticker_values.csv")
complete_meta %>% write_csv("data/ticker_meta.csv")
```
